---
title: "NBA Midterm"
author: "Amber Duevel"
date: "2/28/2022"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```


```{r library, include=FALSE}
Sys.setenv("VROOM_CONNECTION_SIZE"=131072 * 2)
library(nbastatR)
library(shiny)
library(tibble)
library(tidyverse)
library(ggplot2)
library(grid)
library(jpeg)
library(RCurl)
library(gganimate)
library(tidyverse)
library(gapminder)
library(gifski)
library(lubridate)
library(maps)
library(mapdata)
library(transformr)
library(ggthemes)
library(dplyr)
library(patchwork)
library(png)
library(directlabels)
```

```{r shot graph, echo=FALSE, warning=FALSE}
court<-rasterGrob(readJPEG("Use_nba_court.jpg"),width=unit(1,"npc"), height=unit(1,"npc")) #creating court and sizing correctly

logo <- rasterGrob(readJPEG("logo.jpg"),width=unit(1.7, "npc"), height=unit(1.5, "npc")) #creating and sizing the logo of the team

shotData2013 <- teams_shots(teams="Minnesota Timberwolves", seasons = 2013)
shotData2014 <- teams_shots(teams="Minnesota Timberwolves", seasons = 2014)
shotData2015 <- teams_shots(teams="Minnesota Timberwolves", seasons = 2015)
shotData2016 <- teams_shots(teams="Minnesota Timberwolves", seasons = 2016)
shotData2017 <- teams_shots(teams="Minnesota Timberwolves", seasons = 2017)
shotData2018 <- teams_shots(teams="Minnesota Timberwolves", seasons = 2018)
shotData2019 <- teams_shots(teams="Minnesota Timberwolves", seasons = 2019)
shotData2020 <- teams_shots(teams="Minnesota Timberwolves", seasons = 2020)
#gathering each season data for the Minnesota Timberwolves

animation_shotData <- bind_rows(shotData2013, shotData2014, shotData2015, shotData2016, shotData2017, shotData2018, shotData2019, shotData2020) #putting all the seasons together into one data frame

Type_of_shot <- ggplot(animation_shotData, aes(x=locationX, y=locationY)) + 
      annotation_custom(court, xmin=-250, xmax=250, ymin= -50, ymax=420) +
      geom_point(aes(color = typeEvent), alpha=0.25) +
      xlim(-250, 250) +
      ylim(-50, 420) +
      coord_fixed() +
      theme(plot.title = element_text(hjust = 0.5,face="bold"))+
      labs(x="Horizontal Length (1/10th ft)", y="Vertical Length (1/10th)", color= "Missed or Made Shot")+
      scale_color_manual(values=c('#1E6194','#79C019'))+
      annotation_custom(logo, xmin=180, xmax=250, ymin= 380, ymax=430)
      


 animation_plot <- Type_of_shot + 
    transition_states(yearSeason)+
    enter_fade()+
    exit_shrink()+
    view_follow()+
    transition_time(yearSeason)+
    ggtitle("Shot Chart\n Season: {frame_time}", label= unique(animation_shotData$nameTeam))
 
animation_plot
animate(animation_plot, nframes= 8, duration = 10)
```
Findings:  
-Three points shots are missed more compared to two point shots.  
-Layups are made more than they are missed.  
-In the 2020 season, there are less shots inside the three point line compared to previous seasons.  
-Very few half court or far away shots are taken but almost all are missed shots.
-A lot of shots are taken as layups or near the hoop.  

Explanation:  
  When looking at the graph there is more green displayed at the three point line compared to blue. This means more shots are missed than made at this line. There is also very few shots that are taken far from the three point line (close to half court). Almost all of these points are green for each season. We can view these points as buzzer beater shots that are shot at the last seconds of the game. Most of these shots are missed which can inform us that these are difficult shots to make. The team has a very low probability of making this type of shot. Each season has a new spot where more shots are taken. In 2020 season, we can see that there are less shots taken inside the three point line compared to other seasons. In every season there is a lot of blue at the net or near it. This means that a lot of layups are made each season. This can inform us that making layups can be seen as an easier shot to make. Or this is the Timberwolves strongest type of shot. Other information we can gather from this is rebounds made. If a player misses a shot, another player can rebound it and make the shot. This could be a strength for the Timberwolves and that's why there is more blue points near the net. I also found it interesting how throughout every season the graph shows more green points than blue. This means that the team has more missed shots overall. Knowing this information we can infer that the Timberwolves might not be a strong shooting team. 

```{r win graph, echo=FALSE, warning=FALSE}
win <- win_probability(game_ids = c(21700002), filter_non_plays = T, nest_data = FALSE, return_message = TRUE)#gathering win probability data
view(win)

plot<-ggplot(win, aes(x=minuteGame, y=pctWinProbHome))+
      geom_line(aes(y=pctWinProbHome, color=nameTeamHome), alpha=.8)+
      geom_line(aes(y=pctWinProbAway,color=nameTeamAway), alpha=.8)+
      labs(x="Amount of Minutes into Game", y="Probability the Home Team will win", color="Team")+
      ggtitle("Probabilty of Winning Vs Minutes into Game")+
      scale_color_manual(values = c('#006BB8','#D31145'))

animation_plot <- plot + transition_reveal(minuteGame)+
  view_follow(fixed_y=TRUE)

animation_plot

```
  
For this graph I used a new data set which was the probability of winning the game. I compared two teams, Golden State Warriors and Houston Rockets. As the game went on the graph shows the probability of each team winning. The graph fluctuates throughout the game and there isn't a set increase or decrease for either team. At some points during the game the lines move towards fifty percent which could mean the game is close. Once there is only a couple minutes left in the game the lines move toward 80 percent and 20 percent. Due to the lines not ending at 100 and 0 percent it means the score difference wasn't a huge gap.



